### 2017-06-13 manuscript/manuscript.Rmd
Work on intro--motivation.

Although in the abstract reproducibility seems easy, it has in large parts failed in practice. One study found that X% of statistical claims in papers could not be reproduced. Some reasons for this included the original data not being available; personnel being changed leading to materials not being organized; not clear which data was used for analysis; not clear which analyses were done.

To target these key threats to reproducibility, many authors have offered that research scientists could do better to curate the key assets of their work: the data and code used to analyze the data. Recognizing that data in the modern workflow is usually in electronic form, and the procedures to analyze it almost exclusively being run on computers, the empirical sciences have recognized certain aspects of computer science to help with this respect. One idea gaining traction in the research community is the idea of version control systems (VCS).

## Aims of the current manuscript

We introduce the popular VCS Git. Illustrate its use in the scientific workflow with a hypothetical example project. Provide commands / instructions for Apple and Windows users. Show how to use it with minimal learning curve with the R Studio IDE. [TODO]

@@

VCSs are computer programs designed for "tracking changes in computer files and coordinating work on those files among multiple people" (Wikipedia). In this tutorial, we focus on the most widely used (within science) VCS, called Git. It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (?? Windows), Git is not a standalone program for navigating files and folders on a computer. Instead, it adds functionality to the OSs existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI) / integrated development environment (IDE)--that allow taking snapshots of the project and its files, and distributing the work across multiple computers and users. To begin using Git, users must first download and install the Git software on their computers. Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others), and is used by major software developers such as Microsoft, Google, and Facebook. It can be freely downloaded at <https://git-scm.com/>.

>>>>

<!-- TODO work on motivating the use of Git: why is keeping track of materials an issue? Is it an issue? How does VCS help? -->

Although reproducibility might at first appear easy, it has in large parts failed in practice. For example, one survey of leading Psychology journals found that 34-58% of articles published between 1985 and 2013 had at least one inconsistently reported *p*-value [@nuijten_prevalence_2015]. Upon recalculation, the authors couldn't obtain the same *p*-values from the reported test statistics, and therefore these 34-58% of articles were, to some degree, non-reproducible.

Why, then, do so many research products fall short of the basic scientific standard of reproducibility? We suggest one reason is the poor level of organization and curation of all the research assets (data and code, usually) that play a part in creating the research product. Researchers usually have no formal methods or accepted gold standards for curating their research assets, and collaborative workflows, especially, are difficult to manage in a reproducible manner. When the complexity and amount of materials related to a project increases, it might be difficult to link analysis files to the correct data files, and track the evolution of both code and data throughout the research cycle.

In fact, the effort to organize and curate materials has been cited as an important reason for why data is so rarely shared in Psychology. In one study, researchers asked authors of 141 Psychology articles (with a total of 249 studies) to share their data for reanalysis [@wicherts_poor_2006]. 73% of the authors refused to share their data, and Wicherts et al. suggested that they did so because it takes considerable effort to clean, document and organize datasets. How could we make organizing and curating materials and data easier, so as to improve the reproducibility of our science?

## Aims of the article

Fortunately for the empirical sciences, challenges related to organizing and curating materials across time, space, and personnel have been solved to a high standard in computer science with Version Control Systems (VCS). In the remainder of this article, we introduce a popular VCS called Git, and illustrate its use in the scientific workflow with a hypothetical example project. In the tutorial below, we provide show how to use the Git VCS to curate your research materials by using text commands from the computers command line, and with a Graphical User Interface called R Studio.

@@

VCSs are computer programs designed for "tracking changes in computer files and coordinating work on those files among multiple people" (Wikipedia). In this tutorial, we focus on the most widely used (within science) VCS, called Git.

<!-- Insert illustration of VCS about here. -->

It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (?? Windows), Git is not a standalone program for navigating files and folders on a computer. Instead, it adds functionality to the OSs existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI) / integrated development environment (IDE)--that allow taking snapshots of the project and its files, and distributing the work across multiple computers and users. To begin using Git, users must first download and install the Git software on their computers. Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others), and is used by major software developers such as Microsoft, Google, and Facebook. It can be freely downloaded at <https://git-scm.com/>.
